Mathematical Aspects of the Capital Market

### 2015-07-15 papers/Number1/Applications.tex
Correct the justification of the ratio of successive eigenvalues converging to 1.

Figure \ref{fig:EigenRatio} shows the ratio of successive eigenvalues
of the matrix $a_{np}^{-2}RR'$, where the rows of matrix $R$ comprise
the rank-transformed sequences of a number of selected time series
in the S\&P 500 index.
\begin{figure}[htb!]
  \centering
  \subfigure[]{
    \includegraphics[scale=0.40]{EigenRatioSP500_100_shown.pdf}
  }
  \subfigure[]{
    \includegraphics[scale=0.40]{EigenRatioSP500_200_shown.pdf}
  }
  \subfigure[]{
    \includegraphics[scale=0.40]{EigenRatioSP500_300_shown.pdf}
  }
  \subfigure[]{
    \includegraphics[scale=0.40]{EigenRatioSP500_400_shown.pdf}
  }
  \caption{$\lambda_{(i+1)} / \lambda_{(i)}$ versus $i$}
  \label{fig:EigenRatio}
\end{figure}
It is seen that, as $i \to \infty$, $\lambda_{(i+1)} / \lambda_{(i)}
\to 1$. Indeed, as argued in previous sections, $ \lambda_{(i)} \to
\Gamma_k^{-2/\alpha} v_{(l)}$.  Now consider $\lambda_{(i+1)}$. One
possibility is $\lambda_{(i+1)} \to \Gamma_{k+1}^{-2/\alpha}
v_{(l)}$. In this case
\begin{eqnarray*}
  {\lambda_{i+1} \over \lambda_{i}} &\to& \left(
    \Gamma_{i} \over \Gamma_{i+1} \right)^{2/\alpha}
  \overset{d}{=} U_{1,i} \overset{a.s.}{\longrightarrow} 1\text{, as }i
  \to \infty
\end{eqnarray*}
where $U_{a,b}$ denotes the $a$-th upper order statistic in a uniform
sample of size $b$.
So $ \lambda_{i+1} \to \Gamma_{k+1}^{-2/\alpha} v_{(l)}$ can be untrue
only if $\exists m\exists l'(m,l') \neq (1, l)$ such that $ \lambda_{i+1} \to
\Gamma_{k+m}^{-2/\alpha} v_{(l')}$ and
\begin{equation}
  \label{eq:X_A}
  \Gamma_{k+m}^{-2/\alpha} v_{(l')} \overset{a.s.}{\longrightarrow}
  \Gamma_{k}^{-2/\alpha} v_{(l)}
\end{equation}
Suppose $m > 1$.
\[
\left(
  \Gamma_{k} \over \Gamma_{k+m}
\right)^{2/\alpha} \overset{d}{=} (U_{m,k+m-1})^{2/\alpha}
\]
The variance of $(U_{m,k+m-1})^{2/\alpha}$ remains finite regardless
of $k$ and $m$ (\textcolor{red}{to be proven}), so the almost sure
convergence in \eqref{eq:X_A} cannot happen.

>>>>

Figure \ref{fig:EigenRatio} shows the ratio of successive eigenvalues
of the matrix $a_{np}^{-2}RR'$, where the rows of matrix $R$ comprise
the rank-transformed sequences of a number of selected time series
in the S\&P 500 index.
\begin{figure}[htb!]
  \centering
  \subfigure[]{
    \includegraphics[scale=0.40]{EigenRatioSP500_100_shown.pdf}
  }
  \subfigure[]{
    \includegraphics[scale=0.40]{EigenRatioSP500_200_shown.pdf}
  }
  \subfigure[]{
    \includegraphics[scale=0.40]{EigenRatioSP500_300_shown.pdf}
  }
  \subfigure[]{
    \includegraphics[scale=0.40]{EigenRatioSP500_400_shown.pdf}
  }
  \caption{$\lambda_{(i+1)} / \lambda_{(i)}$ versus $i$}
  \label{fig:EigenRatio}
\end{figure}
It is seen that, as $i \to \infty$, $\lambda_{(i+1)} / \lambda_{(i)}
\to 1$. This is in fact expected based on the theory explained in the previous sections. Suppose the sequences are iid and serial correlations are absent, i.e. the matrix $M=H H'$ has only one eigen value. Then it is clear ${\lambda_{i+1} \over \lambda_i}=\left({\Gamma_i \over \Gamma_{i+1}}\right)^{2/\alpha}=\left(1 - {E_{i+1} \over E_1 + E_2 + \dots E_{i+1}}\right)^{2/\alpha} \to 1$, where $E_k, k=1,2,\dots,i+1$ are iid. exponential rv.
% Indeed, as argued in previous sections, $ \lambda_{(i)} \to
% \Gamma_k^{-2/\alpha} v_{(l)}$.  Now consider $\lambda_{(i+1)}$. One
% possibility is $\lambda_{(i+1)} \to \Gamma_{k+1}^{-2/\alpha}
% v_{(l)}$. In this case
% \begin{eqnarray*}
%   {\lambda_{i+1} \over \lambda_{i}} &\to& \left(
%     \Gamma_{i} \over \Gamma_{i+1} \right)^{2/\alpha}
%   \overset{d}{=} U_{1,i} \overset{a.s.}{\longrightarrow} 1\text{, as }i
%   \to \infty
% \end{eqnarray*}
% where $U_{a,b}$ denotes the $a$-th upper order statistic in a uniform
% sample of size $b$.
% So $ \lambda_{i+1} \to \Gamma_{k+1}^{-2/\alpha} v_{(l)}$ can be untrue
% only if $\exists m\exists l'(m,l') \neq (1, l)$ such that $ \lambda_{i+1} \to
% \Gamma_{k+m}^{-2/\alpha} v_{(l')}$ and
% \begin{equation}
%   \label{eq:X_A}
%   \Gamma_{k+m}^{-2/\alpha} v_{(l')} \overset{a.s.}{\longrightarrow}
%   \Gamma_{k}^{-2/\alpha} v_{(l)}
% \end{equation}
% Suppose $m > 1$.
% \[
% \left(
%   \Gamma_{k} \over \Gamma_{k+m}
% \right)^{2/\alpha} \overset{d}{=} (U_{m,k+m-1})^{2/\alpha}
% \]
% The variance of $(U_{m,k+m-1})^{2/\alpha}$ remains finite regardless
% of $k$ and $m$ (\textcolor{red}{to be proven}), so the almost sure
% convergence in \eqref{eq:X_A} cannot happen.
